Stability in human interaction networks: primitive typology of vertex, prominence of measures and time activity statistics

### 2015-04-29 paper.tex
arrumado layout
arrumacoes espalhadas (ateh related work)

\begin{abstract}
 This article reports interaction networks stability by means of three quantitative criteria: activity distribution in time and among participants; a sound classification of vertices in peripheral, intermediary and hub sectors; the combination of basic measures into principal components with greater variance. 
 We analyzed the temporal activity and topology evolution of networks in four email lists by considering window sizes from 50 to 10,000 messages, which were made to slide to generate snapshots of the network along a timeline. Activity in terms of seconds and minutes exhibit an uniform pattern, while hours, days and months exhibit stable concentrations.
 Participant activity follows the expected distribution of scale-free networks. 
 We compare these networks to Erd\"os-R\'enyi networks in order to assign members to three distinct sectors, namely hubs, intermediary and periphery.
 Approximately 5\% of the vertices are hubs, 15-35\% are intermediary and the remainder belongs periphery. 
 The metrics that most contribute to data dispersion were found to be centrality-related (degree, strength and betweenness), followed by symmetry-related, and then clustering coefficient. 
 The results also embrace a sketch of a physics-based typology of agents, with proper social and psychological speculations.
 Because the network properties reported did not depend on the email list and were stable over time, and because observed structure is in accordance with expectations driven from literature for human interaction networks, we believe that the properties observed are also present in general human interaction networks.
 Current unfoldings include governance and accountability proposals and implementations, anthropological physics experiments, audiovisual reconnaissance, and the report of quantitative differences of textual production as connectivity of agents changes.
\end{abstract}

@@

The evolution of human interaction networks was addressed with a community focus, where the direction of edges was not taken into account~\cite{barabasiEvo}. Two topologically different networks emerged from human interaction networks, depending on the frequency of interactions, which can either be a generalized power law or an exponential connectivity distribution~\cite{barabasiTopologicalEv}. In email list networks, scale-free properties were reported with $\alpha=1$~\cite{bird} (as are web browsing and library loans~\cite{barabasiHumanDyn}), and different linguistic traces were related to weak and strong ties~\cite{GMANE2}.

@@

\begin{table}
	\caption{Example of activity percentages along the days of the week. Higher activity was observed during weekdays, with a decrease of activity on weekends of at least one third and two thirds in extreme cases.}
\begin{center}
    \begin{tabular}{ | l |  c | c | c | c | c |   c | c |}
        \hline
        & Mon & Tue & Wed & Thu & Fri & Sat & Sun  \\ \hline
	\input{tables/tabWeekdays}
    \end{tabular}
\end{center}
\label{tab:win}
\end{table}

\begin{table}
	\caption{No significant variation of activity in the days along the month was observed. One cannot point much more than a - probably not statistically relevant - tendency of first and second weeks to be more active. The most important trait seems to be homogeneity.}
	\footnotesize
	\input{tables/tabMonthdaysCPP}
\label{tab:min}
\end{table}

@@

\FloatBarrier

@@

The first principal component is an average of centrality metrics: degrees, strengths and betweenness centrality. Therefore, all of these centrality measurements are equally important for characterizing the networks. On one hand, the relevance of all centrality metrics is not surprising since they may be highly correlated. The degree and strength, for instance, are highly correlated, with Spearman correlation coefficient $\in [0.95,1]$ and Pearson coefficient $\in [0.85,1)$ for $ws>1000$.
On the other hand, each measure relates to a different participation characteristic, and their equal relevance is noticeable.
The clustering coefficient is presented in almost perfect orthogonality to centrality measures.

@@

\begin{figure*} 
   \centering
        \includegraphics[width=.6\textwidth,height=10cm]{figs/im13PCAPLOT_}
	\caption{Prevalence of the symmetry-related metrics over clustering for data dispersion, which was an important result from the PCA analysis, together with stability of component formation, and can be observed in Table~\ref{tab:pcain}.
		The first plot shows degree versus clustering coefficient.
		The second plot is similar, but the first component is an average of centrality metrics.
		The third plot resulted from applying PCA with symmetry-related metrics, where greater dispersion in the second component can be observed among all sectors.
		In this case, the second component is representative of symmetry measurements of vertex interactions and clustering coefficient is only relevant for the third component.
		Greater dispersion suggests symmetry-related metrics as being more powerful for characterizing interaction networks than clustering coefficient, especially for hubs and intermediary vertices.
		This figure was done with a snapshot of the LAU list in a window size of $ws = 1000$ messages.
		The general layout is recurrent and consistent with the literature: most connected vertices have low clusterization while higher clusterization is gradually more incident as the number of connections is lowered.
		Similar structures were observed in all window sizes $ws\;\in\;[500,10000]$ and for networks of other email lists, which points to a common relationship between the measures of degree, strength and betweenness centrality, the measures of symmetry and the measure of clustering coefficient described in Section~\ref{measures}.
    }
    \label{fig:sym}
\end{figure*}

>>>>

\begin{abstract}
 This article reports interaction networks stability by means of three quantitative criteria: activity distribution in time and among participants; a sound classification of vertices in peripheral, intermediary and hub sectors; the combination of basic measures into principal components with greater variance. 
 We analyzed the temporal activity and topology evolution of networks in four email lists by considering window sizes from 50 to 10,000 messages, which were made to slide to generate snapshots of the network along a timeline. Activity in terms of seconds and minutes exhibit an uniform pattern, while hours, days and months exhibit stable concentrations.
 Participant activity follows the expected distribution of scale-free networks. 
 We compare these networks to Erd\"os-R\'enyi networks in order to assign members to three distinct sectors, namely hubs, intermediary and periphery.
 Approximately 5\% of the vertices are hubs, 15-35\% are intermediary and the remainder belongs periphery. 
 The metrics that most contribute to the dispersion of participants in the topological measures space were found to be centrality-related (degree, strength and betweenness), followed by symmetry-related, and then clustering coefficient. 
 The results also embrace a sketch of a physics-based typology of agents.
 Because the network properties reported did not depend on the email list and were stable over time, and because observed structure is in accordance with expectations driven from literature for human interaction networks, we believe that the properties observed are also present in general human interaction networks.
 Current unfoldings include governance and accountability proposals and implementations, anthropological physics experiments, audiovisual reconnaissance, and the report of quantitative differences of textual production as connectivity of agents changes.
\end{abstract}

@@

Network types have been reported relating number of participants, intermittence of them and network longevity~\cite{barabasiEvo}. Such results are confluent and endorsed by results of Section~\ref{sec:pty}. 
Two topologically different networks are reported to emerge from human interaction networks, depending on the frequency of interactions, which can either be a generalized power law or an exponential connectivity distribution~\cite{barabasiTopologicalEv}. In email list networks, scale-free properties were reported with $\alpha=1$~\cite{bird} (as are web browsing and library loans~\cite{barabasiHumanDyn}), and different linguistic traces were related to weak and strong ties~\cite{GMANE2}.

@@

\FloatBarrier
\begin{table}
	\caption{Example of activity percentages along the days of the week. Higher activity was observed during weekdays, with a decrease of activity on weekends of at least one third and two thirds in extreme cases.}
\begin{center}
    \begin{tabular}{ | l |  c | c | c | c | c |   c | c |}
        \hline
        & Mon & Tue & Wed & Thu & Fri & Sat & Sun  \\ \hline
	\input{tables/tabWeekdays}
    \end{tabular}
\end{center}
\label{tab:win}
\end{table}
\begin{table}
	\caption{No significant variation of activity in the days along the month was observed. One cannot point much more than a - probably not statistically relevant - tendency of first and second weeks to be more active. The most important trait seems to be homogeneity.}
	\footnotesize
	\input{tables/tabMonthdaysCPP}
\label{tab:min}
\end{table}

@@

@@

The first principal component is an average of centrality metrics: degrees, strengths and betweenness centrality. Therefore, all of these centrality measurements are equally important for characterizing the networks. On one hand, the relevance of all centrality metrics is not surprising since they may be highly correlated. The degree and strength, for instance, are highly correlated, with Spearman correlation coefficient $\in [0.95,1]$ and Pearson coefficient $\in [0.85,1)$ for $ws>1000$.
On the other hand, each measure relates to a different participation characteristic, and their equal relevance is noticeable.
The clustering coefficient is presented in almost perfect orthogonality to centrality measures.
\begin{figure*} 
   \centering
        \includegraphics[width=.6\textwidth,height=10cm]{figs/im13PCAPLOT_}
	\caption{Prevalence of the symmetry-related measures over clustering for dispersion of participants in the space of topological measures. This is an important result from the PCA analysis, together with stability of component formation, and can be observed in Table~\ref{tab:pcain}.
		The first plot shows degree versus clustering coefficient.
		The second plot is similar, but the first component is an average of centrality metrics.
		The third plot resulted from applying PCA with symmetry-related metrics, where greater dispersion in the second component can be observed among all sectors.
		In this case, the second component is representative of symmetry measurements of vertex interactions and clustering coefficient is only relevant for the third component.
		Greater dispersion suggests symmetry-related metrics as being more powerful for characterizing interaction networks than clustering coefficient, especially for hubs and intermediary vertices.
		This figure was done with a snapshot of the LAU list in a window size of $ws = 1000$ messages.
		The general layout is recurrent and consistent with the literature: most connected vertices have low clusterization while higher clusterization is gradually more incident as the number of connections is lowered.
		Similar structures were observed in all window sizes $ws\;\in\;[500,10000]$ and for networks of other email lists, which points to a common relationship between the measures of degree, strength and betweenness centrality, the measures of symmetry and the measure of clustering coefficient described in Section~\ref{measures}.
    }
    \label{fig:sym}
\end{figure*}

@@
